---
title: "Welcome"
format: 
  html:
    theme: csafe.css
---

HandwriterRF is designed to assist forensic document examiners by performing a statistical analysis on two handwriting samples of unknown origin. Two hypotheses are considered:

$H_p: \text{The two documents were written by the same writer.}$
$H_d: \text{The two documents were written by different writers.}$

The statistical analysis produces a *score-based likelihood ratio (SLR)*. An SLR greater than one, indicates that the evidence supports $H_p$ over $H_d$, and the larger the SLR, the stronger the support. An SLR less than one, indicates that the evidence supports $H_d$ over $H_p$, and the closer the SLR is to zero, the stronger the support.

# Quick Start

## Installation

HandwriterRF requires R and RStudio IDE.

-   Install R from [POSIT](https://posit.co/download/rstudio-desktop/)
-   Install RStudio IDE from [POSIT](https://posit.co/download/rstudio-desktop/)

Install the handwriterRF R package. Open RStudio, navigate to the console window, and type

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("CSAFE-ISU/handwriterRF")
```

## Compare Two Handwriting Samples

Scan and save two handwriting samples as PNG images. Open RStudio, navigate to the console window, and type:

```{r, eval = FALSE}
library(handwriterRF)

calculate_slr(sample1_path = "path/to/handwriting_sample1.png",
              sample2_path = "path/to/handwriting_sample2.png")
```

The result is a *score-based likelihood ratio (SLR)* that quantifies the strength of evidence in favor of "same writer" or "different writer." 
